Title here

### 2017-01-04 portfolio/text/ms.Rmd
More edits and references

Individuals who rely on the extraction of natural resources for their livelihood
--- fishers, farmers, miners, and loggers, among others --- face high levels of
income variability driven by a mix of environmental, biological, and economic
variability. A key component to managing these industries is identifying how
regulatory actions and individual behaviour affect income variability and by
extension the economic stability and sustainable use of natural resources. For
commercial fisheries, recent work has identified that communities and vessels
that fish a greater diversity of species have less revenue variability than
those that fish a lower diversity. However, it is unclear if these benefits
extend to the level of the individual and how year-to-year changes in
diversification affect revenue and revenue variability. Here we identify three
axes by which fishers in Alaska, one of the world's most productive fishing
regions, can diversify fishing activities. We show that fishing permits with
higher species diversity reduces individual revenue variability and fishing an
additional permit is associated with higher revenue and lower variability.
However, increasing species diversity within the constraints of existing permits
has a fishery-dependent effect on revenue and is usually (`r g2_dens`%
probability) associated with increased revenue uncertainty the following year.
Our results demonstrate that the most effective option for individuals to
decrease revenue variability is to purchase permits to participate in additional
or more diverse fisheries. However, this option is costly and often limited by
regulations such as limited entry and catch share programs. With increasing
environmental variability in a changing climate, it will be particularly
important that individuals who rely on natural resources for their livelihood
have effective strategies to reduce income variability.

<!-- TODO: fit something about specialization over time into the abstract.
Thus, over the past 30 years as the Alaskan commercial fishing fleet has become
increasingly specialized, -->

@@

In contrast to the clear across-strategy effects of species diversity on revenue
and revenue variability, the effect of changing fished species diversity for an
individual *within* the constraints of a given set of permits is more variable
(Fig. \ref{fig:within}). After accounting for fishing effort, specializing tends
to be associated with increased revenue but also increased variability (Fig.
\ref{fig:within}; Figs S\ref{fig:fe}, S\ref{fig:re}, S\ref{fig:noeffort}). On
average, there is a ~0.5% increase in the year-to-year revenue ratio per 1%
decrease in species diversity, and across strategies weighted equally, `r
g2_dens`% of the probability density supports a positive relationship between
specialization and revenue.

@@

For individuals deriving livelihoods from natural resources, theory suggests
that adopting strategies that exploit a diversity of assets provides substantial
benefits over strategies with lower diversity [REFS]. However, in practice, the
ability of individuals to effectively diversify is limited by regulation and
market forces. Here we document the effect of individual-level changes in
species diversity on revenue and revenue variability across three scales. We
find that (1) individuals with sets of permits that allow high levels of species
diversity tend to have lower revenue variability. (2) Adding a fisheries permit
usually increases revenue and decreases revenue variability. However, (3) within
the constraints of a permit and accounting for effort, specializing tends to
increase revenue while both specializing and generalizing tend to correlate with
greater revenue variability with considerable differences amongst fisheries.
Thus, the most effective mechanism by which species diversity can lower revenue
variability for individual fishers --- switching or adding fishing permits ---
is costly, limited by regulations, and consequently unavailable to many
individuals.

@@

Our analysis provides a synthetic view of the effects of species diversity on
individuals over 30 years in one of the most productive fishing regions in the
world. However, this large temporal and geographical scope potentially obscures
time or regional-variation in the consequences of diversification on revenue and
revenue variability. We expect future research examining diversification over
finer scales will uncover additional interesting relationships. Furthermore, in
addressing the overall pattern, we have averaged out several major discrete
events, such as the Exxon Valdez oil spill in 1989, that had significant
consequences for the ecological and social systems in Alaska [@paine1996]. In
addition to revenue or income variability, it is important to understand the
benefits of diversification in the context of rare and extreme disturbances
which occur in natural systems worldwide [REF].

<!--
Worldwide, fisheries confront the challenge of sustainable management (). While
many fisheries are well managed and productive, many are also rebuilding from
periods of over-exploitation (Worm et al. Science 2009). A common tactic to aid
in fisheries rebuilding is to modify regulations to reduce the unintended
harvest rebuilding stocks while allowing the continued harvest of healthy stocks
(need REF). Such bycatch reduction is an important goal for both ecological
(Crowder and Murawski 1998, Lewison et al. 2004) and economic (Curtis and Hicks
2000; Chakravorty and Nemoto 2001) reasons. Cumulatively, though, regulatory
changes for bycatch reduction as well as other reasons gradually lead to
fisheries that target one or a few species and individual fishers wishing to
participate in multiple fisheries must own (or purchase) multiple permits. Thus
a general consequence of fisheries regulations that are shown to improve
sustainability of fish stocks and avoid overfishing - including improved
assessments (Hilborn and Ovando 2014) and individual quota systems (Costello et
al. 2008 ) - may ultimately restrict individual diversification or at least make
developing a diverse fishing portfolio more costly. Add Alaska example of
IFQs...?
-->

@@

TODO: finish this:
External market forces may interact in concert with management regulations to
promote specialization for fishers or other individuals dependent on natural
resources. With increasing environmental variability and extreme events in
a changing climate (REF), it is particularly important that individuals who rely
on natural resources as a primary source of income have strategies to reduce
risk. Though variability in individual revenue may not be actively incorporated
into many management decisions, analyses like ours help to understand the range
of diversification strategies available and how their benefits may change over
time.

@@

Our models focused on species diversity as a possible predictor of revenue and
revenue variability while controlling for effort. For consistency with previous
analyses [@vedeld2007; @illukpitiya2008; @debela2012; @kasperski2013;
sethi2014], we calculated effective species diversity for permit holder $i$ and
time (year) $t$ as the inverse of Simpson's diversity index $\lambda$
[@simpson1949; @jost2006] (also referred to in economics as the
Herfindahl-Hirschman Index) weighted by revenue $R$:

@@

\noindent where $S_{i,t}$ is the log ratio of species diversity from
year-to-year, $\log \left(\mathrm{div}_{i,t}/\mathrm{div}_{i,t-1}\right)$,
$D_{i,t}$ is the log ratio of days fished from year-to-year, $\log
\left(\mathrm{days}_{i,t}/\mathrm{days}_{i,t-1}\right)$, and $I_{i,t}$ is an
indicator variable that takes a value $0$ if $\mathrm{div}_{i,t} \ge
\mathrm{div}_{i,t-1}$ (species diversity increasing; generalizing) and takes
a value $1$ if $\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}$ (species diversity
decreasing; specializing). Estimating a separate coefficient for the effect of
increasing diversity and decreasing diversity allows the costs and benefits of
these changes to be asymmetric.

>>>>

Individuals relying on the extraction of natural resources for their livelihood
--- fishers, farmers, miners, and loggers, among others --- face high levels of
income variability driven by a mix of environmental, biological, management, and
economic variability. A key component to managing these industries is
identifying how regulatory actions and individual behaviour affect income
variability and by extension the economic stability and sustainable use of
natural resources. For commercial fisheries, recent work has identified that
communities and vessels fishing a greater diversity of species have less revenue
variability than those fishing a lower diversity. However, it is unclear if
these benefits extend to individual fishers and how year-to-year changes in
diversification affect revenue and revenue variability. Here, we identify three
axes by which fishers in Alaska, one of the world's most productive fishing
regions, can diversify fishing activities. We show that, despite increasing
specialization over the last 30 years, fishing permits with higher species
diversity reduces individual revenue variability and fishing an additional
permit is associated with higher revenue and lower variability. However,
increasing species diversity within the constraints of existing permits has
a fishery-dependent effect on revenue and is usually (`r g2_dens`% probability)
associated with increased revenue uncertainty the following year. Our results
demonstrate that the most effective option for individuals to decrease revenue
variability is to purchase permits to participate in additional or more diverse
fisheries. However, this option is expensive, often limited by regulations such
as catch share programs, and consequently unavailable to many individuals. With
increasing environmental variability in a changing climate, it will be
particularly important that individuals relying on natural resources for their
livelihood have effective strategies to reduce income variability.

<!-- TODO: take home is the management context is key and controls show that
it's more complicated. TODO mentioned that specializing tends to be profitable
-->

@@

In contrast to the clear across-strategy effects of species diversity on revenue
and revenue variability, the effect of changing fished species diversity for an
individual *within* the constraints of a given set of permits is more variable
(Fig. \ref{fig:within}). After accounting for fishing effort and changes in
annual average revenue for fishers with the same set of permits, specializing
tends to be associated with increased revenue but also increased variability
(Fig. \ref{fig:within}; Figs S\ref{fig:fe}, S\ref{fig:re}, S\ref{fig:noeffort}).
On average, there is a ~0.5% increase in the year-to-year revenue ratio per 1%
decrease in species diversity, and across strategies weighted equally, `r
g2_dens`% of the probability density supports a positive relationship between
specialization and revenue.

@@

For individuals deriving livelihoods from natural resources, theory suggests
that adopting strategies that exploit a diversity of assets provides substantial
benefits over strategies with lower diversity [@baldursson1997; @hilborn2001;
@edwards2004]. However, in practice, the ability of individuals to effectively
diversify is limited by regulation and market forces. Here we document the
effect of individual-level changes in species diversity on revenue and revenue
variability across three scales. We find that (1) individuals with sets of
permits that allow high levels of species diversity tend to have lower revenue
variability. (2) Adding a fisheries permit usually increases revenue and
decreases revenue variability. However, (3) within the constraints of a permit
and accounting for effort, specializing tends to increase revenue while both
specializing and generalizing tend to correlate with greater revenue variability
with considerable differences amongst fisheries. Thus, the most effective
mechanism by which species diversity can lower revenue variability for
individual fishers --- switching or adding fishing permits --- is expensive,
limited by regulations, and consequently unavailable to many individuals.

@@

Our analysis provides a synthetic view of the effects of species diversity on
individuals over 30 years in one of the most productive fishing regions in the
world. However, this large temporal and geographical scope potentially obscures
time or regional-variation in the consequences of diversification on revenue and
revenue variability. We expect future research examining diversification over
finer scales will uncover additional interesting relationships. Furthermore, in
addressing the overall pattern, we have averaged out several major discrete
events, such as the Exxon Valdez oil spill in 1989, that had significant
consequences for the ecological and social systems in Alaska [@paine1996]. In
addition to revenue or income variability, it is important to understand the
benefits of diversification in the context of rare and extreme disturbances
which occur in natural systems worldwide [@anderson2016].

<!--
Worldwide, fisheries confront the challenge of sustainable management (). While
many fisheries are well managed and productive, many are also rebuilding from
periods of over-exploitation (Worm et al. Science 2009). A common tactic to aid
in fisheries rebuilding is to modify regulations to reduce the unintended
harvest rebuilding stocks while allowing the continued harvest of healthy stocks
(need REF). Such bycatch reduction is an important goal for both ecological
(Crowder and Murawski 1998, Lewison et al. 2004) and economic (Curtis and Hicks
2000; Chakravorty and Nemoto 2001) reasons. Cumulatively, though, regulatory
changes for bycatch reduction as well as other reasons gradually lead to
fisheries that target one or a few species and individual fishers wishing to
participate in multiple fisheries must own (or purchase) multiple permits. Thus
a general consequence of fisheries regulations that are shown to improve
sustainability of fish stocks and avoid overfishing - including improved
assessments (Hilborn and Ovando 2014) and individual quota systems (Costello et
al. 2008 ) - may ultimately restrict individual diversification or at least make
developing a diverse fishing portfolio more expensive. Add Alaska example of
IFQs...?
-->

@@

TODO: finish this:
External market forces may interact in concert with management regulations to
promote specialization for fishers or other individuals dependent on natural
resources. With increasing environmental variability and extreme events in
a changing climate [@easterling2000], it is particularly important that
individuals who rely on natural resources as a primary source of income have
strategies to reduce risk. Though variability in individual revenue may not be
actively incorporated into many management decisions, analyses like ours help to
understand the range of diversification strategies available and how their
benefits may change over time.

@@

Our models focused on species diversity as a possible predictor of revenue and
revenue variability while controlling for effort. For consistency with previous
analyses [@vedeld2007; @illukpitiya2008; @debela2012; @kasperski2013;
sethi2014], we calculated effective species diversity for permit holder $i$ and
time (year) $t$ as the inverse of Simpson's diversity index $\lambda$
[@simpson1949; @jost2006] (also referred to in economics as the
Herfindahl-Hirschman Index [@hirschman1964]) weighted by revenue $R$:

@@

\noindent where $S_{i,t}$ is the log ratio of species diversity from
year-to-year, $\log \left(\mathrm{div}_{i,t}/\mathrm{div}_{i,t-1}\right)$,
$D_{i,t}$ is the log ratio of days fished from year-to-year, $\log
\left(\mathrm{days}_{i,t}/\mathrm{days}_{i,t-1}\right)$, and $I_{i,t}$ is an
indicator variable that takes a value $0$ if $\mathrm{div}_{i,t} \ge
\mathrm{div}_{i,t-1}$ (species diversity increasing; generalizing) and takes
a value $1$ if $\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}$ (species diversity
decreasing; specializing). Estimating a separate coefficient for the effect of
increasing diversity and decreasing diversity allows the costs and benefits of
these changes to be asymmetric (Fig. S\ref{fig:resid}).
